Algorithm explanation

### 2013-08-20 article/algo.tex

@@

The Horn clause solving algorithm receives an input problem in Horn
graph. A \textbf{Horn graph} is a labelled directed graph $G=(V,E,\varphi)$.
\begin{itemize}
\item $V$ is a union of two disjoint sets of vertices; Horn term
  vertices $V_T$ and arrow vertices $V_\Rightarrow$. Each vertex $u
  \in V_\Rightarrow$ has eactly one outgoing edge to some $v \in V_T$,
  denoted as $succ(u)$.
\item $E$ is a set of labelled edges, which are expressed in the form
  $(u,v,\theta)$ where $\theta$ is a finite map from variables to
  variables.
\item $\varphi: V_T \rightarrow L$ is a labelling function for Horn
  term vertices, where L is a set of elements of a predicate variable
  in the form $P(x_1, \ldots, x_{\mathrm{arity}(P)})$ or a linear
  inequality $a_1 x_1 + \cdots + a_n x_n \leq b$.
\end{itemize}
There are two kinds of edges in a Horn graph;
\begin{itemize}
\item for $u \in V_T, v \in V_\Rightarrow$, an edge takes a form
  $(u,v,\theta)$, written $\edgel{u}{\theta}{v}$, or,
\item for $u \in V_\Rightarrow, v \in V_T$, an edge takes a form
  $(u,v,\emptyset)$, written $\edge{u}{v}$, where the mapping is
  empty.
\end{itemize}

@@

We implicitly assume each $\mathcal{HC}_v$ is universally quantified
by its all free variables $\textsc{FV}(\mathcal{HC}_v)$.  Without loss
of generality, we limit linear expression labels on $V_T$ vertices
with incoming edges to either $0 \leq -1$, meaning $\bot$, or $0 \leq
0$, meaning $\top$.

The solution of $G$ is a map $\rho$ from predicate variables to linear
predicates of the form $\lambda x_1, \cdots ,x_n. \psi $ (where $\psi$
is a formula constructed from
\begin{align*}
\psi ::= & a_1 x_1 + \cdots + a_n x_n \leq b \mid \\
& \psi \wedge \psi \mid \psi \vee \psi \mid \top \mid \bot
\end{align*}
), where $\rho[G]$ is a tautology. A solution is \textit{simple} if
$\rho[P]$ is of the form $a_1 x_1 + \cdots + a_n x_n \leq b$ for every
$P$ in $\mathrm{dom}(\rho)$.

@@

We consider below a Horn graph that is acyclic.

>>>>

\subsection{Example}

@@

Our Horn clause solving algorithm receives an input problem in Horn
graph. A \textbf{Horn graph} is a labelled directed acyclic graph $G=(V,E,\varphi)$.
\begin{itemize}
\item $V$ is a union of two disjoint sets of vertices; Horn term
  vertices $V_T$ and arrow vertices $V_\Rightarrow$. Each vertex $u
  \in V_\Rightarrow$ has eactly one outgoing edge to some $v \in V_T$,
  denoted as $succ(u)$.
\item $E$ is a set of labelled edges, which are expressed in the form
  $(u,v,\theta)$ where $\theta$ is a finite map from variables to
  variables.
\item $\varphi: V_T \rightarrow L$ is a labelling function for Horn
  term vertices, where L is a set of elements of a predicate variable
  in the form $P(x_1, \ldots, x_{\mathrm{arity}(P)})$ or a linear
  inequality $a_1 x_1 + \cdots + a_n x_n \leq b$.
\end{itemize}
There are two kinds of edges in a Horn graph;
\begin{itemize}
\item for $u \in V_T, v \in V_\Rightarrow$, an edge takes a form
  $(u,v,\theta)$, written $\edgel{u}{\theta}{v}$, or,
\item for $u \in V_\Rightarrow, v \in V_T$, an edge takes a form
  $(u,v,\emptyset)$, written $\edge{u}{v}$, where the mapping is
  empty.
\end{itemize}

@@

We implicitly assume each $\mathcal{HC}_v$ is universally quantified
by its all free variables $\textsc{FV}(\mathcal{HC}_v)$.  Without loss
of generality, we limit linear expression labels on $V_T$ vertices
with incoming edges to $1 \leq 0$, meaning $\bot$.

The solution of $G$ is a map $\rho$ from predicate variables to linear
predicates of the form $\lambda x_1, \cdots ,x_n. \psi $ (where $\psi$
is a formula constructed from
\begin{align*}
\psi ::= & a_1 x_1 + \cdots + a_n x_n + b \leq 0 \mid \\
& \psi \wedge \psi \mid \psi \vee \psi \mid \top \mid \bot
\end{align*}
), where $\rho[G]$ is a tautology. A solution is \textit{simple} if
$\rho[P]$ is of the form $a_1 x_1 + \cdots + a_n x_n + b \leq 0$ for every
$P$ in $\mathrm{dom}(\rho)$.

@@

We first propose an algorithm to compute a simple solution for the given
$G$ if one exists. Later we add modifications to the algorithm if a
simple solution does not exist.
Our algorithm works in a constructive manner. List 1 shows the
pseudo-code of our algorithm.
